---
title: "Assignment B4"
author: "Hannah Bobst"
date: "December 7, 2023"
output: 
  html_document:
    theme: cerulean
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE}
# Load packages
library(tidyverse)
library(janeaustenr)
library(stopwords)
```


# Exercise 1
This exercise creates a bar graph of the 20 most common words in the *emma* data set from the *janeaustenr* package. 

### Preparing the data
The data must be prepared before it may be plotted. 
First, each word is made into a single element of the data set. 
The list format is removed so that the data set is a single vector, and all words are converted to be completely lower case. 
The blank observations and punctuation are then removed.
Afterward, all the stop words from the *stopwords* package are removed, and this subset of data is converted to a data frame (for plotting). 
A summary table is then created which summarizes the amount of times each word appears.
This table is sorted from most the common to least common words. 
Finally, the table is truncated so that only the 20 most common words are included.

```{r message=FALSE}
# Create lowercase subset of "emma" data set without stop punctuation and blank entries
emma_subset <- emma %>%
  str_split(pattern=" ") %>%
  unlist() %>%
  tolower() %>%
  str_subset(pattern="") %>%
  str_remove_all(pattern="[.:;,?()_!&]") %>%
  str_remove_all(pattern="\"") %>%
  str_remove_all(pattern="--")

# Subset "emma" further without stopwords (from "stopwords" package)
emma_subset <- emma_subset[(emma_subset %in% stopwords()) == FALSE] 

# Make "emma_subset" into a data frame
emma_subset_df <- data.frame("words" = emma_subset)

# Summarize counts of words and order largest to smallest
emma_counts <- emma_subset_df %>%
  count(words) %>%
  arrange(desc(n))

# Select the 20 most common words for plotting
emma_counts_20 <- emma_counts[1:20,]
```

### Plotting the data
A bar plot is created for the 20 most common words in the *emma* data set. 

```{r message=FALSE}
# Plot the 20 most common words
emma_counts_20 %>%
  ggplot(aes(x = reorder(words, n), y = n)) + 
  geom_col(fill = "darksalmon") + 
  coord_flip() + 
  guides(fill = "none") + 
  geom_text(aes(label=n), hjust = 1.2, colour="white") + 
  ggtitle("20 Most Common Words in Emma") + 
  xlab("Words") + 
  ylab("Word Count")
```

