---
title: "Assignment B4"
author: "Hannah Bobst"
date: "December 7, 2023"
output: 
  html_document:
    theme: cerulean
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE}
# Load packages
library(tidyverse)
library(janeaustenr)
library(stopwords)
```


# Exercise 1
This exercise creates a bar graph of the 20 most common words in the *emma* data set from the *janeaustenr* package. 

### Preparing the data
The data must be prepared before it may be plotted. 
First, each word is made into a single element of the data set. 
The list format is removed so that the data set is a single vector, and all words are converted to be completely lower case. 
The blank observations and punctuation are then removed.
Afterward, all the stop words from the *stopwords* package are removed, and this subset of data is converted to a data frame (for plotting). 
A summary table is then created which summarizes the amount of times each word appears.
This table is sorted from most the common to least common words. 
Finally, the table is truncated so that only the 20 most common words are included.

```{r message=FALSE}
# Create lowercase subset of "emma" data set without stop punctuation and blank entries
emma_subset <- emma %>%
  str_split(pattern=" ") %>%
  unlist() %>%
  tolower() %>%
  str_subset(pattern="") %>%
  str_remove_all(pattern="[.:;,?()_!&]") %>%
  str_remove_all(pattern="\"") %>%
  str_remove_all(pattern="--")

# Subset "emma" further without stopwords (from "stopwords" package)
emma_subset <- emma_subset[(emma_subset %in% stopwords()) == FALSE] 

# Make "emma_subset" into a data frame
emma_subset_df <- data.frame("words" = emma_subset)

# Summarize counts of words and order largest to smallest
emma_counts <- emma_subset_df %>%
  count(words) %>%
  arrange(desc(n))

# Select the 20 most common words for plotting
emma_counts_20 <- emma_counts[1:20,]
```

### Plotting the data
A bar plot is created for the 20 most common words in the *emma* data set. 

```{r message=FALSE}
# Plot the 20 most common words
emma_counts_20 %>%
  ggplot(aes(x = reorder(words, n), y = n)) + 
  geom_col(fill = "darksalmon") + 
  coord_flip() + 
  guides(fill = "none") + 
  geom_text(aes(label=n), hjust = 1.2, colour="white") + 
  ggtitle("20 Most Common Words in Emma") + 
  xlab("Words") + 
  ylab("Word Count")
```

# Exercise 2

### Function
A function *pig_latin_mod* is created which converts a vector of character data to a modified version of Pig Latin. 

This modified Pig Latin has the following rearrangement rules: 

* For words that end with a vowel, all of the last vowels (until the first consonant) are moved to the beginning of the word. 

* For words that end with a consonant, all of the last consonants (until the last vowel) are moved to the beginning of the word 

In addition, this modified Pig Latin has the following addition rule: 

* Add "ya" to the beginning of each word (after the rearrangement rule has been performed)

For example, the word *about* becomes *yatabou*. 
This is because *t* has been moved from the end to the beginning, and *ya* has been added to the beginning. 
```{r message=FALSE}
#' @title Convert words to modified Pig Latin 
#'
#' @description pig_latin_mod takes a vector of character data (one word per observation) and converts each observation to a modified version of Pig Latin. 
#'
#' @param data A vector. 
#' @return A tibble summarizing the maximum values of variable by groups.
#' 
#' @import tidyverse
#' @export
pig_latin_mod <- function(data){
  # Check if input data is a vector
  if(!is.vector(data)){
    # If data isn't a vector, stop function with an error 
    stop("Input data must be a vector.")
  }
  # Check if the vector has character data 
  if(!is.character(data)){
    # If the vector doesn't have character data, stop function with an error
    stop("Input vector must have character data.")
  }
  
  # Function to detect if a letter is a vowel 
  is_vowel <- function(x) x %in% c("a", "e", "i", "o", "u", "A", "E", "I", "O", "U")
  # Function to detect if a letter is a consonant
  is_consonant <- function(x) !(x %in% c("a", "e", "i", "o", "u", "A", "E", "I", "O", "U"))
  
  # Map each letter in function 
  map_vec(data, function(x){
    # Split current word by letter
    data_letters <- str_split(x, "")[[1]]
    
    # Find letter at the end of the current word
    end_letter <- str_sub(x, start=-1)
    
    # Find last vowel index for the current word
    last_vowel <- detect_index(data_letters, is_vowel, .dir="backward")
    # Find last consonant index for the current word
    last_con <- detect_index(data_letters, is_consonant, .dir="backward")
  
    # If the last letter is a consonant 
    if(is_consonant(end_letter) == TRUE){
      # Move letters (until last vowel) from the end to the beginning and add "ya" to the beginning
      str_c("ya", str_sub(x, start=last_vowel+1), str_sub(x, end=last_vowel))
    } else{
      # Move letters (until last consonant) from the end to the beginning and add "ya" to beginning
      str_c("ya", str_sub(x, start=last_con+1), str_sub(x, end=last_con))
    }
  })
  
}
```


### Examples


### Tests
